---
title: "homework10"
author: "Nathaniel Wigdahl"
github repo: https://github.com/nwigdahl44/practice4/tree/main/Homework%2010
date: "4/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
```

```{r import the data sets}
accidents = read.csv("accident.csv")
persons = read.csv("person.csv")
```

### Plot accidents from Latitude and Longitude
```{r plot accident data}
cont_accident <- accidents%>% filter(LONGITUD < 750, LATITUDE < 50, LATITUDE > 24)
states <- map_data("state")
states %>% ggplot(aes(x = long, y = lat)) + geom_path(aes(group = group)) + 
  geom_point(aes(x= LONGITUD, y = LATITUDE), data = cont_accident)


```


## Part One
#### Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?
```{r Accidents by day}
accidents %>% group_by(DAY_WEEK) %>% summarise(total_acc = n())
temp_graph <- accidents %>% group_by(DAY_WEEK) %>% summarise(total_acc = n()) %>% 
  mutate(DAY_WEEK = as.factor(DAY_WEEK))
temp_graph %>% ggplot(aes(x = DAY_WEEK, y = total_acc, fill=DAY_WEEK)) + geom_col()
```

From this basic histogram we can conclude that accidents tend to happen on and around the weekend, with the lowest rates in the middle of the week.
    
#### What time of the day do accidents happen (use variable HOUR)?
```{r}
accidents %>% filter(HOUR != 99) %>% group_by(HOUR) %>% summarise(total_acc = n()) %>%  ggplot(aes(x = HOUR, y = total_acc)) + geom_col()
```

The primary time range accidents occur is from 3-11 pm, early evening early night.

#### What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?
```{r}
accidents %>% filter(DRUNK_DR > 0) %>% summarise(n = n())
```

There were 8720 accidents with at least one drunk driver.

## Part Two

#### Connect to the person table. Identify drivers (PER_TYP == 1, see fars manual ) and subset on them.\
```{r}
 drivers <-  persons %>% filter(PER_TYP == 1) #We only want the drivers. 
 drivers <- drivers %>% select(ST_CASE, HOUR, SEX)  # Select pertinent data for the upcoming join
 drivers <- drivers %>% filter(SEX == 1 | SEX == 2) %>% filter(HOUR != 99)
```  
#### Join accident and driver table (work out which variable(s) to use)
```{r}
 acc_sub <- accidents %>% select(ST_CASE, DAY_WEEK)  #Select columns we want in our joined table

acc_driver <- left_join(drivers, acc_sub, by = "ST_CASE")
```
    
#### Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!
```{r}
#Tally the numbers
graph1 <- acc_driver %>% group_by(DAY_WEEK, SEX) %>% summarise(n = n())
graph2 <- acc_driver %>% group_by(HOUR, SEX) %>% summarise(n = n())

#Visualize results.
graph1 %>% spread(key = SEX, value = n) %>% 
  ggplot() + geom_col(aes(x = DAY_WEEK, y = `1`, fill = 'Men')) + geom_col(aes(x = DAY_WEEK, y = `2`, fill = 'Women'))

graph2 %>% spread(key = SEX, value = n) %>% 
  ggplot() + geom_col(aes(x = HOUR, y = `1`, fill = 'Men')) + geom_col(aes(x = HOUR, y = `2`, fill = 'Women'))
```

### Github repo: https://github.com/nwigdahl44/practice4/tree/main/Homework%2010

